---
title: 'Homework Week 15'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework Week 15

## 1)
Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.
$$( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )$$
```{r}
x <- c( 5.6, 6.3, 7, 7.7, 8.4 )
y <- c( 8.8, 12.4, 14.8, 18.2, 20.8 )
lm_dat <- lm( y ~ x )
plot( x,y ); abline( lm_dat$coefficients, col ='blue' )
```
```{r}
paste( 'The equation for the lm best-fit line: y= ', round(summary( lm_dat )$coefficients[2,1],2), '*x +', summary( lm_dat )$coefficients[1,1])
```




## 2)
Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form $( x, y, z )$. Separate multiple points with a comma.
$$f(x,y) = 24x-6xy^2 -8y^3$$

Find the first partial derivatives $f_x$ and $f_y$
$$f_x(x,y) = 24-6y^2$$
$$f_y(x,y) = 12xy - 24y^2$$
The critical points satisfy the fist partial derivatives set equal to zero:
$$24-6y^2=0 \rightarrow$$
$$24=6y^2 \rightarrow$$
$$y = \sqrt{\frac{24}{6}} = \pm2$$
$$12xy-24y^2=0$$
solving for $y=2$:
$$12x(2)-24(2)^2 = 24x - 96 = x - 4 \rightarrow x = 4$$
solving for $y=-2$:
$$12x(-2)-24(-2)^2 = -24x - 96 = -x - 4 \rightarrow x = -4$$
Therefore, the critical points are (2,4) & (2,-4)

Find the second order partial derivatives $f_{xx}(x,y)$, $f_{yy}(x,y)$ & $f_{xy}(x,y)$
$$f_{xx}(x,y) = 0$$
$$f_{yy}(x,y) = 12x-48y$$
$$f_{xy}(x,y) = -12y$$

Now to use the equation $D = f_{xx}(a,b)f_{yy}(a,b) - f_{xy}^2(a,b)$ for each critical point $(a,b)$:

1. For point (2,4): $D = (0)(12*2-48*4) - (-12*4)^2 \rightarrow$ negative and $f_{xx}(2,4)=0$ therefore this is a saddle point
2. For point (2,-4): $D = (0)(12*2-48*4) - (-12*(-4))^2 \rightarrow$ negative and $f_{xx}(2,-4)=0$ therefore this is a saddle point



## 3)
A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the
"house" brand and $40 + 11x - 23y$ units of the "name" brand.

1. **Step 1** Find the revenue function $R(x,y)$.

$$R( x,y ) = x(81-21x+17y) + y(40+11x-23y)$$

2. **Step 2** What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?
```{r}
x <- 2.3
y <- 4.1
R <- x*( 81-21*x+17*y ) + y*( 40+11*x-23*y )
paste0( 'Revenue = $', round( R, 2 ) )
```



## 4)
A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

It is given that:
$$\mbox{Total units produced} = x + y = 96$$
Can solve for x, where $x= 96-y$ and plug this expression into the Cost equation:
$$C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700 \rightarrow$$
$$= \frac{1}{6}(96-y)^2 + \frac{1}{6}y^2 + 7(96-y) + 25y + 700 \rightarrow$$
$$= \frac{1}{3}y^2 - 14y + 2908$$
Taking the 1st derivative:
$$C'(y) = \frac{2}{3}y -14 $$
To find the critical point:
$$\frac{2}{3}y -14 = 0 \rightarrow y = 21$$
Substituting back into the 1st equation:
$$x + 21 = 96 \rightarrow x = 75$$
To minimize cost, the LA location should produce 75 units while the Denver location should produce 21.

## 5)
Evaluate the double integral on the given region.

$$\int \int (e^{8x+3y})dA;  R: 2 \leq x \leq 4 \mbox{ and } 2 \le y \le 4$$
Write your answer in exact form without decimals.

First find the intgral with respect to x:
$$\int_2^4 e^{8x+3y}dx = \frac{1}{8}(e^{3y+32}-e^{3y+16})$$
Now integrate with respect to y:
$$\int_2^4 \frac{1}{8}(e^{3y+32}-e^{3y+16}) dy = \frac{e^{44} -e^{38} -e^{28} + e^{22}}{24}$$



<br><br><br>