---
title: 'DATA605: Assignment 2'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
# Trace / Determinant / Factorization
## Problem Set #1

(1) Show that $A^T A \neq AA^T$ in general. (Proof and demonstration.)

&nbsp;&nbsp;&nbsp;&nbsp;If matrix $A$ has dimentions $(m , n)$ and $A^T$ is $A$'s transpose, then the products $A^TA$ and $AA^T$ will have the dimensions $(n,n)$ and $(m,m)$ respectively. Therefore, if $m \neq n$ it follows that $A^T A \neq AA^T$  

&nbsp;&nbsp;&nbsp;&nbsp;This is demonstrated here:

```{r}
#Generally, A^TA !=AA^T
#Demonstration
A <- matrix( c( 1,2,3,4,1,1,1,1,2,0,1,2 ), nrow=3 )
print( A )
print( t( A ) )
```

```{r error=TRUE}
left_side = t( A ) %*% ( A )
right_side = ( A ) %*% t( A )

paste( 'size of AtA: ', dim( left_side )[1], 'x', dim( left_side )[2], '    size of AAt: ', dim( right_side )[1], 'x', dim( right_side )[2], sep="" )

identical( dim( left_side ), dim( right_side ) )
```


(2) For a special type of square matrix A, we get $A^T A = AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

&nbsp;&nbsp;&nbsp;&nbsp;Square matices with symmetry across the diagonal are a special case where $A^T A = AA^T$ is true.

```{r}
#Conditions where A^TA == AA^T
A = matrix( c( 1,2,3,2,1,2,3,2,1 ), nrow=3 )
left_side = t( A ) %*% ( A )
right_side = ( A ) %*% t( A )
print( left_side )
print( right_side )
identical( left_side , right_side )
```


## Problem Set #2

&nbsp;&nbsp;&nbsp;&nbsp;Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.  
&nbsp;&nbsp;&nbsp;&nbsp;Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code.

```{r}
#R function to factorize a square matrix A into LU or LDU
#Doolittle Algorithm for LU Decomposition as described: www.https://www.geeksforgeeks.org/doolittle-algorithm-lu-decomposition/

doolittle_LUdecomp <- function( A, n ) {
  #decompose square matrix A with (n x n) dimentions into lower and upper triagular factors
  
  #initialize lower and upper
  lower = matrix( 0L, nrow = n, ncol = n )
  upper = matrix( 0L, nrow = n, ncol = n )
  
  #LU decomposition
  for (a in 1:n){
    
    #upper triagle
    for (b in a:n){
      #sum of L(a,c)*U(c,b)
      sum <- 0
      for (c in 1:a){
        sum <- sum + ( lower[a,c] * upper[c,b] )
      }
      #eval U(a,b)
      upper[a,b] <- A[a,b] - sum
    }
    #lower triangle
    for (b in a:n){
      if ( a ==b ){
        lower[a,a] <- 1 #diagonal =1
      } else {
        #sum of L(b,c)*U(c,a)
        sum <- 0
        for (c in 1:a){
          sum <- sum + ( lower[b,c] * upper[c,a] )
        }
        #eval L(b,a)
        lower[b,a] <- ( A[b,a] - sum) / upper[a,a]
      }
    }
    
  }
  return( list( 'U' = upper, 'L' = lower ) )
}
```

```{r}
A = matrix( c( 2,-4,-4,-1,6,-2,-2,3,8 ), nrow=3 )
result <- doolittle_LUdecomp( A, dim( A )[1] )
result
```

