---
title: "DATA605: Assignment 4"
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
# Inverse of a Matrix / Single Value Decomposition
## Problem Set #1

Given a 3 × 2 matrix A:
$$A = \begin{bmatrix}
1 & 2 & 3 \\
-1 & 0 & 4
\end{bmatrix}$$

write code in R to compute $\mathbf{X} = \mathbf{AA^T}$ and $\mathbf{Y} = \mathbf{A^TA}$. Then, compute the eigenvalues
and eigenvectors of $\mathbf{X}$ and $\mathbf{Y}$ using the built-in commands in R.

```{r}
A = matrix( c( 1,-1,2,0,3,4 ), nrow = 2, ncol = 3 )
print( A )
```
```{r}
X = A %*% t( A )
Y = t( A ) %*% A
print( X )
print( Y )
```
<br>  
Compute the eigenvalues and eigenvectors of X and Y using the built-in commans in R.  

```{r}
Xeig <- eigen( X )
Yeig <- eigen( Y )
print( 'eigenvalues of X:' )
print( Xeig['values'] )
print( 'eigenvalues of Y:' )
print( Yeig['values'] )
print( 'eigenvectors of X:' )
print( Xeig['vectors'] )
print( 'eigenvectors of Y:' )
print( Yeig['vectors'] )
```
<br>
Compute the left-singular, singular values, and right-singular vectors of A using the svd command:
```{r}
components = svd( A )
print( 'singular values of A' )
print( components['d'] )
print( 'left singular vectors of A' )
print( components['u'] )
print( 'right singular vectors of A' )
print( components['v'] )
```
<br>
Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y. In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A:
```{r}
#will use the asrenal library's comparedf
library( arsenal )
#recast as dataframes and use comparedf
#get the summary and assign to a variables
cmp_Xu <- summary( comparedf( as.data.frame( Xeig['vectors'] ), as.data.frame( components['u'] ) ) )
cmp_Yv <- summary( comparedf( as.data.frame( Yeig['vectors'] ), as.data.frame( components['v'] ) ) )
#we are only interested in the differences between the values in the dataframes
numdiffs_Xu <- n.diffs( cmp_Xu ) #determines how many differences were found between values
numdiffs_Yv <- n.diffs( cmp_Yv )
paste( 'Num differences between the eigenvectors of X and the left singular vectors of A: ', numdiffs_Xu)
paste( 'Num differences between the eigenvectors of Y and the right singular vectors of A: ', numdiffs_Yv)
#ref 1 for more info
```
<br>
```{r}
#comparing the eigenvalues of X, Y and the singular values of A
numdigits <- 10
test1 <- round( Xeig['values'][[1]][1], numdigits ) == round( Yeig['values'][[1]][1], numdigits )
test2 <- round( Xeig['values'][[1]][2], numdigits ) == round( Yeig['values'][[1]][2], numdigits )

#comparing X to the squares of A
test3 <- round( Xeig['values'][[1]][1], numdigits ) == round( components['d'][[1]][1]^2, numdigits )
test4 <- round( Xeig['values'][[1]][2], numdigits ) == round( components['d'][[1]][2]^2, numdigits )

if ( test1 == TRUE & test2 == TRUE )
  print( 'The 2 nonzero eigenvalues of X & Y are equal' )

if ( test3 == TRUE & test4 == TRUE )
  paste( 'The 2 eigenvalues of X are equal to the square of the singular values of A (it follows that this holds for the non zero eigenvalues of Y as well)' )
```

## Problem Set #2

write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors.  

The Inverse Matrix will be calculated with these steps (ref#2):  

1. calculate the matrix of minors 
2. find the matrix of cofactors
3. find the adjugate
4. multiply by 1/Determinant

```{r}
#functionalize finding the co-factor
findCofactor = function( M ){
  cofactor_M <- M
  stopifnot( length( unique( dim( M ) ) )==1 ) #make sure M is square
  sizeM <- dim( M )[1]
  for(i in 1:sizeM ){
    for( j in 1:sizeM ){
      cofactor_M[i,j] <- ( det(M[-i,-j]) * (-1)^(i+j) ) 
      #STEP1 the first term (det(M[-i,-j])) finds the matrix of minors
      #STEP2 the second term ((-1)^(i+j)) alternates +/- for matrix of cofactors
    }                                               
  }
  return( cofactor_M )
}


myinverse = function(M){
  cofactor <- findCofactor( M )
  T_cofactor <- t( cofactor ) #STEP3 take the transpose of the cofactor to find the adjugate (adjoint)
  det_M <- det(M)
  M_inverse <- T_cofactor/det_M #STEP4 multiply by 1/det to find the inverse
  return( M_inverse )
}
```

```{r}
C <- matrix(c(1,5,5,3,3,1,7,5,4), nrow=3, ncol=3, byrow=TRUE)
invC <- myinverse( C )
print( C )
print( invC )
```
```{r}
#damage control
damage_control <- C %*% invC
print( round(damage_control, digits = 14 ) )
```


## References
1. [The `comparedf` function](https://cran.r-project.org/web/packages/arsenal/vignettes/comparedf.html#extract-differences)
2. [Inverse of a Matrix using Minors, Cofactors and Adjugate](https://www.mathsisfun.com/algebra/matrix-inverse-minors-cofactors-adjugate.html)