---
title: 'DATA605: Assignment 3'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
# Eigenvalues / Eigenvectors
## Problem Set #1

(1) What is the rank of the matrix A?

$$A = \begin{bmatrix}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & -1 \\
5 & 4 & -2 & -3
\end{bmatrix}$$

The rank is 4, because there are no redundant rows. this can be varified:
```{r}
library( Matrix )
A = matrix( c( 1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,-1,-3 ), nrow = 4, ncol = 4 )
rankMatrix( A )
```

(2) Given an `mxn` matrix where `m > n`, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

If `A` is an `mxn` matrix:
$$rank( A_{mxn} ) \leq min( m,n )$$
Therefore, if `m > n`, it follows that the max rank of `A` is `n`

The minimum rank of a non-zero matrix `A` is 1

(3) What is the rank of matrix B?

$$B = \begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2
\end{bmatrix}$$

The rank of `B` is 1 which can be shown using row reduction:

\begin{equation*}
B = \begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2
\end{bmatrix}
\dashrightarrow
(R_3 = R_3 - 2*R_1)
\dashrightarrow
\begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
0 & 0 & 0
\end{bmatrix}
\dashrightarrow
(R_2 = R_2 - 3*R_1)
\dashrightarrow
\begin{bmatrix}
1 & 2 & 1 \\
0 & 0 & 0 \\
0 & 0 & 0
\end{bmatrix}
\end{equation*}

This can also be varified with r:
```{r}
B = matrix( c( 1,3,2,2,6,4,1,3,2 ), nrow = 3, ncol = 3 )
rankMatrix( B )
```

## Problem Set 2

Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your
work. You’ll need to write out the characteristic polynomial and show your solution.

$$
A = \begin{bmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6
\end{bmatrix}
$$

We use the following equations to find the eigenvalues:
$$
det(A - \lambda I) = 0
$$

$$ det
\begin{pmatrix}
1-\lambda & 2 & 3 \\
0 & 4-\lambda & 5 \\
0 & 0 & 6-\lambda
\end{pmatrix}
= 0
$$
Exanding the determinant we get:

$$(1-\lambda )(4-\lambda )(6-\lambda ) + (2)(5)(0) + (3)(0)(0) - (2)(0)(6-\lambda ) - (1-\lambda )(5)(0) - (3)(4-\lambda )(0) = 0$$
Removing the terms with 0 simplifies the result to 

$$(1-\lambda )(4-\lambda )(6-\lambda ) = 0$$
To satisfy this equation, we find the 3 eigenvalues to be:

1. $\lambda = 1$
2. $\lambda = 4$
3. $\lambda = 6$

Now to find the eigenvectors for each eigenvalue by solving $N( A-\lambda_i I)$

(1) $\lambda = 1$

$$
\left[
\begin{array}{ccc|c}
1-1 & 2 & 3 & 0 \\
0 & 4-1 & 5 & 0 \\
0 & 0 & 6-1 & 0 \\
\end{array}
\right]
\longrightarrow
\left[
\begin{array}{ccc|c}
0 & 2 & 3 & 0 \\
0 & 3 & 5 & 0 \\
0 & 0 & 5 & 0 \\
\end{array}
\right]
$$

Now to solve with gaussian elimination
$$
\left[
\begin{array}{ccc|c}
0 & 2 & 3 & 0 \\
0 & 3 & 5 & 0 \\
0 & 0 & 5 & 0 \\
\end{array}
\right]
\rightarrow
(R_1/2)
\rightarrow
\left[
\begin{array}{ccc|c}
0 & 1 & \frac{3}{2} & 0 \\
0 & 3 & 5 & 0 \\
0 & 0 & 5 & 0 \\
\end{array}
\right]
\rightarrow
((R_2-3)/\frac{1}{2})
\rightarrow
\left[
\begin{array}{ccc|c}
0 & 1 & \frac{3}{2} & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 5 & 0 \\
\end{array}
\right]
\rightarrow
(R_3-5*R_2)
\rightarrow
\left[
\begin{array}{ccc|c}
0 & 1 & \frac{3}{2} & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
$$
This gives up
Var2 = 0 (row 1 equation), Var3 = 0 (row 2 equation) and Var1 = Var1
so the first eigenvector =
$$v_1 = \begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}$$

(2) $\lambda = 4$

$$
\left[
\begin{array}{ccc|c}
1-4 & 2 & 3 & 0 \\
0 & 4-4 & 5 & 0 \\
0 & 0 & 6-4 & 0 \\
\end{array}
\right]
\longrightarrow
\left[
\begin{array}{ccc|c}
-3 & 2 & 3 & 0 \\
0 & 0 & 5 & 0 \\
0 & 0 & 2 & 0 \\
\end{array}
\right]
$$

Now to solve with gaussian elimination
$$
\left[
\begin{array}{ccc|c}
-3 & 2 & 3 & 0 \\
0 & 0 & 5 & 0 \\
0 & 0 & 2 & 0 \\
\end{array}
\right]
\rightarrow
(R_1/-3)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{3} & -1 & 0 \\
0 & 0 & 5 & 0 \\
0 & 0 & 2 & 0 \\
\end{array}
\right]
\rightarrow
(R_2/5)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{3} & -1 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 2 & 0 \\
\end{array}
\right]
\rightarrow
(R_3-2*R_2)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{3} & -1 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
\rightarrow
(R_1+*R_2)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{3} & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
$$

This gives up
Var3 = 0 (row 2 equation), Var1 - 2/3*Var2 = 0 (row1 equation), so
Var1 = 2/3Var2
so the first eigenvector =
$$v_2 = \begin{bmatrix}
\frac{2}{3} \\
1 \\
0
\end{bmatrix}$$

(3) $\lambda = 6$

$$
\left[
\begin{array}{ccc|c}
1-6 & 2 & 3 & 0 \\
0 & 4-6 & 5 & 0 \\
0 & 0 & 6-6 & 0 \\
\end{array}
\right]
\longrightarrow
\left[
\begin{array}{ccc|c}
-5 & 2 & 3 & 0 \\
0 & -2 & 5 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
$$

Now to solve with gaussian elimination
$$
\left[
\begin{array}{ccc|c}
-5 & 2 & 3 & 0 \\
0 & -1 & 5 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
\rightarrow
(R_1/-5)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{5} & \frac{-3}{5} & 0 \\
0 & -2 & 5 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
\rightarrow
(R_2/-2)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & \frac{-2}{5} & \frac{-3}{5} & 0 \\
0 & 1 & \frac{-5}{2} & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
\rightarrow
(R_1-\frac{-2}{5} *R_2)
\rightarrow
\left[
\begin{array}{ccc|c}
1 & 0 & \frac{-8}{5} & 0 \\
0 & 1 & \frac{-5}{2} & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
$$

This gives up two equations:
Var1 - 8/5Var3 = 0  -> Var1 = 8/5Var3
Var2 - 5/2Var3 = 0  -> Var2 = 5/2Var3

so the first eigenvector =
$$v_3 = \begin{bmatrix}
\frac{8}{5} \\
\frac{5}{2} \\
1
\end{bmatrix}$$


This can be verified with r eigen():
```{r}
A = matrix( c( 1,0,0,2,4,0,3,5,6 ), nrow = 3, ncol = 3 )
eigen( A )
```



