---
title: 'DATA605: Assignment #6'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# (1)

Let $X_1, X_2, . . . , X_n$ be $n$ mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to $k$. Let $Y$ denote the
minimum of the $X_i$’s. Find the distribution of $Y$.

To find the distribution, start by evaluating the probability where the minimum value of the distributions is 1: 
$$P(Y=1) = 1 - P(Y>1) = 1-\left( \frac{k-1}{k} \right)^n$$
Next, evaluate $P(Y=2)$

$$P(Y=2) = 1 - P(Y>2) -P(Y=1)$$
substitute in the answer above:
$$P(Y=2) = 1 - P(Y>2) - \left( 1-\left( \frac{k-1}{k} \right)^n \right)$$
$P( Y>2 ) = \left( \frac{k-2}{k} \right)^n$, substituting this in to the previous result gives:
$$P(Y=2) = 1 - \left( \frac{k-2}{k} \right)^n - \left( 1-\left( \frac{k-1}{k} \right)^n \right)$$
This can be simplified to
$$P(Y=2) = \frac{(k-2)^n-(k-1)^n}{k^n}$$
This can be generalized as:
$$P(Y=y) = \frac{(k-y)^n-(k-y-1)^n}{k^n}$$

Now to functionalize this expression:
```{r}
Pofy <- function( k,y,n ) {
   return ((k-(y-1))^n - (k-(y))^n)/k^n
}
```
Evaluate an example where:
```{r}
n <- 20.0 #mutually independent random variables
k <- 10.0 #max range of the uniform distribution
```

```{r, message=FALSE}
Pvals <- c()
for ( i in seq( 1,k,1 ) ){
  val <- Pofy( k,i,n )
  Pvals[i] <- val
}

library( ggplot2 )
Xvals <- seq( 1,k,1 )
data <- data.frame( Xvals, Pvals )
ggplot( data, aes( x=Xvals, y=Pvals ) ) +
  geom_line() +
  ggtitle( 'Distribution of Minimums' )
```



# (2)

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).

## (a)
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

```{r}
P_F <- 1/10 #probability of failing in a given year
P_NF <- 1 - P_F #probability of not failing in a given year
E_val <- 1 / P_F #expected value
STD <- sqrt( P_NF / P_F^2 ) #standard deviation
#probability that the machine will not fail for the first 8 years
# P( no failures for 8 years ) = 1 - P( not failing )^8
numY <- 8
P_8yearsNF_geometic <- 1-( 1-P_NF^numY )
#alternatively
P_8yearsNF_geometic <- 1 - pgeom( numY-1, P_F )

answer <- paste( 'Expected Value:', E_val,
                 '\nStandard Deviation:', round( STD, 4 ),
                 '\nProbability of no failures for 8 years:', round( P_8yearsNF_geometic, 4 ) )
cat( answer, sep='\n' )
```


## (b)
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential. 

Exponential Distribution
$$P(x;\lambda ) = \lambda e^{-\lambda x}$$
```{r}
E_val <- 1 / P_F #expected value
STD <- sqrt( 1 / P_F^2 ) #standard deviation
P_8yearsNF_exp <- pexp( numY, P_F, lower.tail = FALSE )

answer <- paste( 'Expected Value:', E_val,
                 '\nStandard Deviation:', round( STD, 4 ),
                 '\nProbability of failure after 8 years (Exponential):', 
                 round( P_8yearsNF_exp, 4 ) )
cat( answer, sep='\n' )
```


## (c)
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years) 

Binomial Distribution
$$\left( \begin{array}{c} n \\ k \end{array} \right) =
\frac{n!}{(n-k)!k!}$$

```{r}
E_val <- numY*P_F #expected value
STD <- sqrt( numY *P_F*P_NF ) #standard deviation
k <- 0 #dbinom will use choose() to compute coefficients
P_8yearsNF_binomial <- dbinom( k, numY, P_F )

answer <- paste( 'Expected Value:', E_val,
                 '\nStandard Deviation:', round( STD, 4 ),
                 '\nProbability of failure after 8 years (Binomial):', 
                 round( P_8yearsNF_binomial, 4 ) )
cat( answer, sep='\n' )
```

## (d)
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

Poisson Distribution:
$$P(x;\lambda)= \frac{e^{-\lambda } * \lambda ^x }{x!}$$
```{r}
lam <- P_F * numY
E_val <- P_F * numY #expected value
STD <- sqrt( P_F * numY ) #standard deviation
P_8yearsNF_poisson <- ppois( 0, lam )

answer <- paste( 'Expected Value:', E_val,
                 '\nStandard Deviation:', round( STD, 4 ),
                 '\nProbability of failure after 8 years (Exponential):', 
                 round( P_8yearsNF_poisson, 4 ) )
cat( answer, sep='\n' )
```

<br><br><br>