---
title: 'Data 605 Homework Wk13'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library( ggplot2 )
```

## #1
Use integration by substitution to solve the integral below.
$$\int 4e^{-7x}dx$$
1. move constant in front
$$\longrightarrow 4 \int e^{-7x}dx$$
2. substitute in $u=-7x$
$$-4\int e^udu\longrightarrow -4\int -\frac{1}{7}e^udu\longrightarrow -\frac{4}{7}\int e^u du$$
3. apply common form for $\int e^udu = e^u$
$$\longrightarrow -\frac{4}{7}e^u$$
4. substitute back in for u
$$\longrightarrow -\frac{4}{7}e^{-7x}$$
Therefore, we find that 
$$\int 4e^{-7x}dx \longrightarrow -\frac{4}{7}e^{-7x}+C$$


## #2
Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt} = -\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function $N( t )$ to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

Find $N(t)$ given $\frac{dN}{dt} = -\frac{3150}{t^4}-220$:

$$\frac{dN}{dt} = -\frac{3150}{t^4}-220 \longrightarrow $$
take the integral of both sides of the equation to find N(t)
$$\int \frac{dN}{dt} = \int \left( -\frac{3150}{t^4}-220 \right)dx \longrightarrow $$
$$\int \frac{dN}{dt} = \int \left( -3150t^{-4}-220 \right)dx \longrightarrow $$
$$N(t) = -\frac{3150 \cdot t^{-3}}{-3}-220t + C \longrightarrow $$
$$N(t) = -\frac{3150 \cdot t^{-3}}{-3}-220t + C \longrightarrow $$
$$N(t) = -1050 \cdot t^{-3}-220t + C \longrightarrow $$
Solve for $C$ given the observation that at $t=1$, $N(t)=6530:
$$6530 = -1050 \cdot (1)^{-3}-220\cdot (1) + C \longrightarrow $$
$$6530 = 830 + C \longrightarrow C = 5700$$


## #3
Find the total area of the red rectangles in the figure below, where the equation of the line is $f ( x ) = 2x - 9$.

Area = 16


## #4
Find the area of the region bounded by the graphs of the given equations:
$$y = x^2 - 2x - 2$$
$$y = x+2$$

Start by visualizing the 2 functions
```{r, message=FALSE}
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
fun.1 <- function(x) x^2 - 2*x -2
fun.2 <- function(x) x +2
p + stat_function(fun = fun.1) + stat_function(fun = fun.2) + xlim(-2,5)
```
great, so we would like to find:
$$Area = \int_{-1}^4 [(x + 2) - (x^2 -2x-2)]dx \rightarrow \int_{-1}^4 [-x^2 + 3x +4 ]dx \rightarrow$$
find the difference between this expression evaluated at 4 and -1:
$$\frac{-x^3}{3} + \frac{3x^2}{2} + 4x$$

```{r}
#ealuated x=4
x <- 4
val4 <- -x^3/3 + 3*x^2/2 + 4*x
x <- -1
valneg1 <- -x^3/3 + 3*x^2/2 + 4*x
Area <- val4 - valneg1
Area
```



## #5
A beauty supply store expects to sell 110 flat irons during the next year. It costs \$3.75 to store one flat iron for one year. There is a fixed cost of \$8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

Consider the sales for a year:
$$110 \mbox{ flat irons} = (\mbox{#orders})\cdot (\mbox{size of each order}) = n\cdot s$$
the store expects to sell 110 total which can be distributed across $x$ number of orders of a size $s$

We want to optimize the number of orders and order size given that we know there are two costs involved: 1) Annual Ordering Costs & 2) Annual Storage Costs

I need to wave a white flag on this problem. I honestly don't believe there s enough information. How can we calculate the Annual Storage Cost if we have no information on how many/how long irons are kept in storage?

## #6
Use integration by parts to solve the integral below.
$$\int ln(9x)\cdot x^6dx$$
apply integration by parts where $u = ln(9x)$ and $dv=x^6dx$, therefore $v=\frac{x^7}{7}$ and $du=\frac{1}{x}dx$ with the equation: $\int udv = uv - \int vdu$:  
$$ln(9x)\cdot \frac{x^7}{7} - \int \frac{x^7}{7}\frac{1}{x}dx \rightarrow 
ln(9x)\cdot \frac{x^7}{7} - \int \frac{x^6}{7}dx \rightarrow$$
$$ln(9x)\cdot \frac{x^7}{7} -  \frac{x^7}{49} \rightarrow ln(9x)\cdot \frac{x^7}{7} -  \frac{x^7}{49} + C$$


## #7
Determine whether $f ( x )$ is a probability density function on the interval $1, e^6$ . If not, determine the value of the definite integral.
$$f(x)=\frac{1}{6x}$$
What is the area under the curve from 1 to $e^6$? If the area = 1, the $f(x)$ is a probability density function....
$$\int_1^{e^6} \frac{1}{6x}dx \rightarrow \frac{1}{6}ln(x)|_1^{e^6}$$
```{r}
x <- 1
val1 <- 1/6*log(x)
x <- exp(6)
valexp6 <- 1/6*log(x)
A = valexp6 - val1
print( A )
```
In conclusion, $f(x) is a probability function over this interval.

<br><br><br>