---
title: 'DATA605: Assignment #6'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# 6.1.6
A die is rolled twice. Let X denote the sum of the two numbers that turn up, and Y the difference of the numbers (specifically, the number on the first roll minus the number on the second). Show that E(XY ) = E(X)E(Y ). Are X and Y independent?

```{r,message=FALSE}
library( dplyr )
#make a dataframe to hold each possible roll
die1 <- c( rep( 1,6 ), rep( 2,6 ), rep( 3,6 ), rep( 4,6 ), rep( 5,6 ), rep( 6,6 ))
die2 <- c( rep( 1:6,6 ) )
combo_df <- data.frame( 'roll1' = die1, 'roll2' = die2 )
combo_df <- combo_df %>% mutate( X = roll1 + roll2, Y = abs(roll1 - roll2), XY = X*Y )
head( combo_df )
```
find the mean of the columns...
```{r}
combo_means <- colMeans( combo_df )
combo_means
```
Does mean( XY ) == mean(X)*mean(Y)?
```{r}
combo_means['XY'] == combo_means['X']*combo_means['Y']
```
This shows that, yes, E(XY ) = E(X)E(Y )

The code below simulates this process and finds the values X, Y and XY
```{r, message=FALSE}

#simulate the double dice roll 'n' times
n <- 10000
roll1 <- sample(1:6, n, replace=TRUE)
roll2 <- sample(1:6, n, replace=TRUE)
#use dplyr methods
die_df <- data.frame( 'roll1' = roll1, 'roll2' = roll2 )
die_df <- die_df %>% mutate( X = roll1 + roll2, Y = abs(roll1 - roll2), XY = X*Y )
head( die_df )
```
Now find the mean of the columns...
```{r}
diemeans <- colMeans( die_df )
diemeans
```

Does mean( XY ) == mean(X)*mean(Y)?
```{r}
diemeans['XY']
diemeans['X']*diemeans['Y']
```
...those simulated values look pretty close

<br><br><br>