---
title: 'DATA605: Discussion Week #10'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# 11.2.14

With the situation in Exercise 13 (the homework problem), consider the strategy such that for $i \lt 4$, Smith bets $min(i, 4 − i)$, and for $i \geq 4$, he bets according to the bold strategy, where $i$ is his current fortune. Find the probability that he gets out of jail using this strategy. How does this probability compare with that obtained for the bold strategy?

Will visualize the system and find the solution using the library `markovchain`.
```{r, message=FALSE}
library(markovchain) #loading the package
```

Define a transition matrix between states in the system. In other words, how Smith's money situation changes as he plays.
```{r}
jailbreak<-matrix(rep(0, 9^2),byrow=TRUE,nrow = 9)
rownames(jailbreak)<-colnames(jailbreak)<-c("$0","$1","$2","$3","$4","$5","$6","$7","$8")
#2 absorption states of the system
jailbreak[1,1] <- 1
jailbreak[9,9] <- 1
for (idx in seq( 2,dim(jailbreak)[1]-1) ) {
  link <- idx-1
  if (idx < 5 ){ #if the state is < 4
    #smith will bet the min( currentstate, 4-currentstate )
    jailbreak[ idx + min( link, 4-link ), idx] = 0.6 #gain 1
    jailbreak[ idx - min( link, 4-link ), idx ] = 0.4 #lose 1
  }
  else{ #if state is >= 4
    #smith will follow bold strategy and bet what he needs to get to '$8' state
    jailbreak[ idx + ( 8 - link ),idx ] = 0.6 #gain 1
    jailbreak[ idx - ( 8 - link ),idx ] = 0.4 #lose 1  
  } 
  
}


jailbreak[2,1] = 0
jailbreak[8,9] = 0
print( jailbreak)
```


```{r}
jailMc<-as(jailbreak, "markovchain")
plot(jailMc, curved=TRUE)
```

```{r}
absp <- absorptionProbabilities(jailMc)
print( absp )
```

Here is what Smith's chances looks like playing with this strategy:  
$P( \mbox{Smith get's out of jail} ) =$ `r absp[2,1]`    
$P( \mbox{Smith stays in jail} ) =$ `r absp[1,1]` 
<br><br><br>