---
title: 'DATA605 Final Project'
subtitle: 'Problem 1'
author: 'Bonnie Cooper'
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Computational Mathematics: Probability

Libraries Used:
```{r, message=FALSE}
library( dplyr )
library( janitor )
```

## Generating Random Numbers

Using 'R', generate a random variable $X$ that has 10,000 random uniform numbers from 1 to $N$, where $N$ can be any number of your choosing greater than or equal to 6. Then generate a random variable $Y$ that has 10,000 random normal numbers with a mean of $\mu = \sigma = (N+1)/2$.

```{r}
# Define N, the max value
N <- 13

# each variable will hold n values where:
n <- 10000
#generate random uniform variable X, with n values from range 1 to N
X <- runif( n, min = 1, max = N )
#generate random normal variable Y, with n values and mean = sigma = (N+1)/2
Y <- rnorm( n, mean = (N+1)/2, sd = (N+1)/2 )

data_df <- data.frame( cbind( X, Y ) )
head( data_df )
```

take a momnt to evaluate the variable:
```{r}
paste( 'length X:', length( X ), ' length Y:',length( Y ) )
print( summary( X ) )
paste( '(N+1)/2 =', (N+1)/2 ) 
print( summary( Y ) )
paste( 'sigma(Y) =', round( sd( Y ), 2 ) )
```

Both $X$ & $Y$ have the expected dimensions and summary statistics


## Probability
Calculate as a minimum the below probabilities a through c.  Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.  Interpret the meaning of all probabilities.

```{r}
#from the problem it is given that:
x <- median( X )
y <- summary( Y )[ 2 ]
res <- paste( 'The median of X, x =', x, '\nThe 1st quartile of Y, y =', y )
cat( res )
```


### a) $P(X \gt x | X \gt y)$

$P(X \gt x | X \gt y) = \frac{P(X \gt x \cap X \gt y)}{P(X \gt y)}$

```{r}
joint <- data_df %>% filter( X > x, Y > y ) 
jointp <- dim( joint )[ 1 ]/ n
marg <- data_df %>% filter( X > y )
margp <- dim( marg )[ 1 ]/ n
condp <- jointp / margp
```

The conditional probability, $P(X \gt x | X \gt y) =$ `r condp`

### b) $P(X \gt x, Y \gt y)$

The joint probability, $P(X \gt x, X \gt y) =$ `r jointp`


### c) $P(X \lt x | X \gt y )$
$P(X \lt x | X \lt y) = \frac{P(X \lt x \cap X \lt y)}{P(X \lt y)}$

```{r}
joint <- data_df %>% filter( X < x, Y < y ) 
jointp <- dim( joint )[ 1 ]/ n
marg <- data_df %>% filter( X < y )
margp <- dim( marg )[ 1 ]/ n
condp <- jointp / margp
```

The conditional probability, $P(X \lt x | X \lt y) =$ `r condp`

### Investigate whether $P(X \gt x \mbox{ & } Y \gt y) = P(X \gt x) P(Y \gt y)$ by building a table and evaluating the marginal and joint probabilities.

The following code builds a joint probability table with marginal probabilities:
```{r}
getjoints <- data_df %>% mutate( jp1 = ( X <= x & Y <= y),
                              jp2 = ( X <= x & Y > y ),
                              jp3 = ( X > x & Y <= y ),
                              jp4 = ( X > x & Y > y ) )
jointps <- colSums( getjoints[,3:6] )/n
jointps <- data.frame( matrix( jointps, ncol=2, byrow=TRUE ) )
colnames( jointps ) <- c( 'P(Y<=y)', 'P(Y>y)' )
rownames( jointps ) <- c( 'P(X<=x)', 'P(X>x)' )

jointps <- jointps %>% mutate( 'Total' = rowSums(.[1:2] ) )
jointps[ 'Total' ,] <- colSums( jointps )
rownames( jointps ) <- c( 'P(X<=x)', 'P(X>x)', 'Total' )
jointps
```

From these values comparisons can be made:  
$P(X \gt x \mbox{ & } Y \gt y) =$ `r jointps[2,2]`  
$P(X \gt x) P(Y \gt y) =$ `r jointps[2,3]` $\cdot$ `r jointps[3,2]` = `r jointps[2,3] * jointps[3,2]`
As a result, we see that the statement $P(X \gt x \mbox{ & } Y \gt y) = P(X \gt x) P(Y \gt y)$ evaluates true for this data; a good indication that the data features are independent.

### Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?

```{r}
getmargins <- data_df %>% mutate( m1 = ( X <= x ),
                              m2 = ( Y <= y ),
                              m3 = ( X > x ),
                              m4 = ( Y > y ) )
margins <- colSums( getmargins[,3:6] )
margins <- matrix( margins, 2, 2, dimnames = list( c( 'X', 'Y' ), c( '<=', '>' ) ) )
margins
```

```{r}
#chi-squared test
test <- chisq.test( margins )
test
```

```{r}
#fisher exact test
fisher.test( margins )
```


The p-values for both tests are very small `p-value < 2.2e-16`, therefore we can reject the null hypothesis and conclude the the two distributions, $X$ & $Y$, are statistically significantly associated. Both the $\chi ^2$ and Fisher's Exact tests are statistical methods of independence between data features. Fisher's test yields an exact result whereas $\chi ^2$ is approximately accurate. However, because this is a very large sample size and $\chi ^2$ accuracy increases with sample size, $\chi ^2$ is more appropriate for this application.






<br><br><br>
